---
title: "HW1"
author: "Ben Schanzer"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

## Question 1

Generate a dataframe with at least ten named individuals. Give them scores on five assignments. Include some missing assignments. Add one column that gives the mean of all their scores, figuring the mean for those with missing assignments based on the number submitted, not the number assigned.

```{r q1}
names = c("Fred", "Ashely", "Owen", "Avery", "Andy", "Andrea", "Emily", "Danny", "Ben", "Conor")

scores <- matrix(c(
  91, 90, 69, 82, 89,
  60, 65, NA, 79, 67,
  75, NA, 95, 72, 84,
  98, 73, 61, NA, 76,
  NA, 80, 70, 93, 62,
  87, 64, 78, 96, 71,
  83, 92, 66, 74, NA,
  68, 77, 100, 63, 88,
  81, NA, 86, 70, 94,
  72, 85, 91, NA, 79
), ncol = 5, byrow = TRUE)

df = as.data.frame(scores)
df = cbind(Student = names, df)

# Row averages ignoring NA values
df$row_avg <- rowMeans(as.matrix(df[,-1]), na.rm = TRUE)

colnames(df) = c("Student", "Test 1", "Test 2", "Test 3", "Test 4", "Test 5", "Average Test Score")

df
```

## Question 2

Using the UKfaculty graph from igraphdata, generate an edgelist and an adjacency matrix. Display the first 10 rows of the edgelist and the upper left 10x10 of the adjacency matrix. Create a separate dataframe listing each node id along with degree, closeness centrality, and betweenness centrality.

```{r q2}
library(igraphdata)
library(igraph)

data("UKfaculty")

A = as_adjacency_matrix(UKfaculty, sparse = FALSE)
E = as_edgelist(UKfaculty)

cat("Adjacency Matrix (top-left 10x10):\n")
print(A[1:10, 1:10])

cat("\nEdge List (first 10 rows):\n")
print(E[1:10, ])

unique_nodes = unique(as.vector(E))
df2 = data.frame(node = unique_nodes)
df2$degree = degree(UKfaculty)[unique_nodes]
df2$closeness = closeness(UKfaculty)[unique_nodes]
df2$betweenness = betweenness(UKfaculty)[unique_nodes]

df2
```

## Question 3

Create an edgelist that includes the connections between you and at least 10 friends or family. Create an igraph object from the edgelist. Plot the graph of your connections, including connections among friends and family.

```{r q3}
edges <- c("Ben", "Jonah",
           "Ben", "Cole",
           "Ben", "Savannah",
           "Ben", "Caden",
           "Ben", "Emily",
           "Ben", "Danny",
           "Ben", "Marty",
           "Ben", "Avery",
           "Ben", "Mom",
           "Jonah", "Cole",
           "Jonah", "Emily",
           "Jonah", "Danny",
           "Jonah", "Mom",
           "Savannah", "Caden",
           "Savannah", "Emily",
           "Savannah", "Danny",
           "Savannah", "Mom",
           "Caden", "Emily",
           "Caden", "Avery",
           "Caden", "Marty",
           "Caden", "Mom",
           "Emily", "Danny",
           "Emily", "Mom",
           "Danny", "Mom",
           "Marty", "Avery",
           "Ben", "Larkin"
)

edge_list <- matrix(edges, ncol = 2, byrow = TRUE)

g = graph_from_edgelist(edge_list, directed = FALSE)
plot(g,
     vertex.size = 25,
     edge.curved = 0.25)
```

## Question 4

**Research Question:**

I am interested in conspiracy theories (CTs) and how they travel and evolve through social networks. Specifically, I want to map networks of conspiracy believers to determine whether there are central nodes — highly connected individuals who act as hubs for spreading new conspiracies through a community. Beyond just the spread of individual theories, I am also curious whether believing in one conspiracy predisposes someone to believe in others. In other words, is someone who believes in Theory A more likely to adopt an unrelated Theory B simply because they are already embedded in a conspiratorial network, or are there other factors — such as political identity, media diet, or personal relationships — that determine which specific theories someone adopts? Understanding this could help researchers identify whether conspiracy belief is driven more by network influence or by individual predisposition.

**Data Acquisition:**

Data could be acquired by scraping public posts and interactions on platforms like Facebook, Twitter/X, or Reddit, using keyword filtering and natural language processing to identify users expressing conspiratorial beliefs. Connections between users could be established through follower relationships, replies, retweets, or shared group memberships. If the goal is to track the flow of a specific conspiracy theory through a network, a directed graph would be appropriate, since we care about who transmitted the theory to whom. This would also allow us to work backwards from widespread belief to identify the original source or sources. If instead the goal is simply to map the community of conspiracy believers and their relationships, an undirected graph would suffice. Weights could serve multiple purposes: on edges, they could represent the frequency or intensity of interaction between two individuals, while node attributes could track how many distinct theories each person believes. Regarding components, the graph would likely have one large connected component encompassing the main conspiracy community, but there may be smaller isolated components representing niche or fringe groups that do not interact with the broader network — for example, a small group focused on a hyper-local conspiracy that has no overlap with mainstream CT communities.

## Question 5

Find the definition of the diameter of a graph. Find the igraph function for measuring diameter. Using the e-mail network as an undirected graph: simplify the network, determine its diameter, and use bootstrapping to determine whether the diameter has a statistically significant difference from a null model of a random graph with the same degree distribution.

**Diameter Definition:** The diameter of a connected (un)directed graph is the farthest distance between any two of its vertices.

```{r q5}
ecomp = as.data.frame(read.table(file = "../Data/email_main_component.txt", sep = "|"))
email_g = graph_from_data_frame(ecomp, directed = FALSE)
email_g = simplify(email_g)

cat("Email network diameter:", diameter(email_g, directed = FALSE), "\n")

# Bootstrapping: compare diameter to null model (random graph with same degree distribution)
d1 = degree(email_g)
t1 = numeric(1000)

for (i in 1:1000) {
  randnet = sample_degseq(d1)
  randnet = simplify(randnet)
  t1[i] = diameter(randnet, directed = FALSE)
}

email_diameter = diameter(email_g, directed = FALSE)
z_score = (email_diameter - mean(t1)) / sd(t1)
cat("Z-score:", z_score, "\n")
```

The high Z-score (>20) indicates that the diameter of the email network is statistically significantly larger than that of a comparable random network with the same degree distribution (Z > 1.96 is significant).
